---
title: "Using an existing classification model"
author: "Vaadia et al"
output:
  html_document: default
  pdf_document: default
editor: visual
---

::: callout-tip
This script will help you using an existing random forest model to classify unobserved behaviors, following the workflow described in Vaadia et al (Figure 2). This script uses example ACC data of unobserved behaviors and the final behavioral classification model created for Griffon Vultures. If you are classifying vulture behavior, you can use our published model. If you are using your own classification model, you will need to upload your own model.

***Important!***

-   Read the instructions before running each chunk to adjust the script to your own dataset.
:::

## Set up the R session

```{r setup, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(moments)
library(tidymodels)
library(ranger)
library(parsnip)
library(caret)
library(zoo)

# Load required functions
# Function to calculate mean amplitude of each bout 
mean_amplitude <- function(x) {
  extreme_points <- which(abs(diff(sign(diff(x)))) == 2) + 1 
  mean(abs(x[extreme_points[-1]] - x[extreme_points[-length(extreme_points)]]))
}

```

## A - Pre-processing sequence

In this section, you will pre-process your unobserved ACC data to create a dataset ready for behavioral classification.

### 1 - Uploading raw ACC data

::: callout-tip
#### ***Important!***

-   If you want to **use your own** **unobserved ACC data**, replace the files inside the folder "Data/Unobserved_Data/" with your own ACC data.
:::

```{r uploading_data}
# Uploading the unobserved raw ACC data
unobs_raw_acc <- list.files(path = "../Data/Unobserved_ACC_data/", 
                            pattern = "*.csv", full.names = T) %>%
  map_df(~read.csv(.)) %>%
  mutate(UTC_datetime = as.POSIXct(UTC_datetime, 
                                   format = c("%Y-%m-%d %H:%M:%S"), 
                                   tz = "UTC"))
# If relevant, remove any rows that do not contain ACC data (for example, rows that have only GPS data)
unobs_raw_acc <- subset(unobs_raw_acc, !(datatype %in% c("GPS", "GPSS")))


```

### 3 - Transform the raw ACC into acceleration values

Each device was calibrated before deployment by measuring the acceleration in all six possible orientations: positive X, negative X, positive Y, negative Y, positive Z, negative Z. We calculated a slope (`slope`) and intercept (`int`) for each axis, representing a device-specific instrument error. Using these values, we can calibrate each device and transform the raw ACC into acceleration values.

For devices without specific error values, we used the average error across all measured devices.

::: callout-tip
#### ***Important!***

-   You need to replace the file "example_calibration.csv" with your own tag calibration file!
:::

```{r calibrating_acc}
# Add calibration file 
calibration <- read.csv("../Data/example_calibration.csv") 

unobs_raw_acc <- left_join(unobs_raw_acc, calibration, by = "device_id")

# If there are no calibration values of a tag, add the mean value for all tags
unobs_raw_acc <- unobs_raw_acc %>% 
  mutate(slopex = ifelse(is.na(slopex), mean(calibration$slopex), slopex),
         intx = ifelse(is.na(intx), mean(calibration$intx), intx),
         slopey = ifelse(is.na(slopey), mean(calibration$slopey), slopey),
         inty = ifelse(is.na(inty), mean(calibration$inty), inty),
         slopez = ifelse(is.na(slopez), mean(calibration$slopez), slopez),
         intz = ifelse(is.na(intz), mean(calibration$intz), intz))

# Transform raw ACC data into acceleration values
unobs_raw_acc <- unobs_raw_acc %>%
  mutate(acc_x = (acc_x - intx) * slopex,
         acc_y = (acc_y - inty) * slopey,
         acc_z = (acc_z - intz) * slopez) %>%
    dplyr::select(-c(intx:slopez))

```

### 4 - Identify distinct bouts

To identify the bouts, there are a few possible scenarios: 1) the device indicates the start of an ACC bout; 2) the ACC is sampling every X minutes, so you can use the time difference between measurements to identify when a new bout started; 3) the ACC is sampling continuously, in which case you can split the continuous data into bouts of X measurements.

In our case (i.e., Ornitela devices), the device identifies the start of a new bout in the column `datatype`, indicated by the variable `SEN_ACC_20Hz_START`. We provide example code for the other two scenarios.

::: callout-tip
#### ***Important!***

-   Use only the code relevant for your data!
:::

```{r identify_bouts}
# ---- 1) The device identifies the start of the bout -----
bout_id <- numeric(nrow(unobs_raw_acc))

j = 0

for(i in 1:nrow(unobs_raw_acc)) {
  if(unobs_raw_acc$datatype[i] == "SEN_ACC_20Hz_START") {
    j = j + 1
    }
  bout_id[i] = j
}

unobs_raw_acc <- unobs_raw_acc %>% 
  add_column(bout_id, .before = 1)


# ---- 2) There is a time difference between bouts ----
# time_threshold <- 300 # add here your relevant time difference, in seconds
# 
# unobs_raw_acc <- unobs_raw_acc %>%
#   group_by(device_id) %>%
#   arrange(UTC_datetime) %>%
#   mutate(time_diff = as.numeric(difftime(UTC_datetime,
#                                          lag(UTC_datetime),
#                                          units = "sec"))) %>%
#   ungroup() %>%
#   mutate(bout_id = cumsum(ifelse(is.na(time_diff) | time_diff >= time_threshold, 1, 0)))


# ---- 3) Continuous sampling ----
# bout_duration <- 5 # add here your relevant bout duration, in seconds
# acc_frequency <- 20 # add here the frequency of ACC collection, in Hz
# 
# bout_length <- bout_duration * acc_frequency
# 
# unobs_raw_acc <- unobs_raw_acc %>%
#   group_by(device_id) %>%
#   arrange(UTC_datetime) %>%
#   mutate(time_diff = as.numeric(difftime(UTC_datetime,
#                                          lag(UTC_datetime),
#                                          units = "sec")),
#          temp_bout_id = (row_number() - 1) %/% bout_length + 1) %>%
#   ungroup() %>%
#   # creates a continuous bout id across the whole dataset, based on the cumulative sum of the changes in the temporary bout id
#   mutate(bout_id = cumsum(temp_bout_id != lag(temp_bout_id, default = 0))) %>%
#   dplyr::select(-temp_bout_id)
# 
# # Making sure that the bouts have continuous data and no gaps (not relevant for the example dataset that had distinct bouts every 300 secs)
# unobs_raw_acc %>%
#   group_by(bout_id) %>%
#   filter(max(time_diff) > 1)

```

### 5 - Exclude incomplete bouts

The bout length is a product of the bout duration (in seconds) and the ACC frequency (in Hz). In the example dataset, the ACC was collected at 20Hz for 5sec, so we expect each bout to have 100 rows.

```{r exclude_incomplete_bouts}
bout_duration <- 5 # add here your relevant bout duration, in seconds
acc_frequency <- 20 # add here the frequency of ACC collection, in Hz

bout_length <- bout_duration * acc_frequency

unobs_raw_acc <- unobs_raw_acc %>%
  add_count(bout_id) %>%
  filter(n == bout_length) %>%
  dplyr::select(-n)

```

### 6 - Extract statistical features

Here we provide a list of statistical features to describe each bout, that are described in Supplementary Table S1. Other features can be added to this function.

```{r extract_stat_features}
stat_feats <- unobs_raw_acc %>%
    group_by(device_id, bout_id) %>%
    summarise(mean_x = mean(acc_x),
              mean_y = mean(acc_y),
              mean_z = mean(acc_z),
              range_x = max(acc_x)-min(acc_x),
              range_y = max(acc_y)-min(acc_y),
              range_z = max(acc_z)-min(acc_z),
              sd_x = sd(acc_x),
              sd_y = sd(acc_y),
              sd_z = sd(acc_z),
              skewness_x = skewness(acc_x),
              skewness_y = skewness(acc_y),
              skewness_z = skewness(acc_z),
              kurtosis_x = kurtosis(acc_x),
              kurtosis_y = kurtosis(acc_y),
              kurtosis_z = kurtosis(acc_z),
              max_x = max(acc_x),
              max_y = max(acc_y),
              max_z = max(acc_z),
              min_x = min(acc_x),
              min_y = min(acc_y),
              min_z = min(acc_z),
              norm_x = sqrt(sum(acc_x^2)),
              norm_y = sqrt(sum(acc_y^2)),
              norm_z = sqrt(sum(acc_z^2)),
              q25_x = quantile(acc_x, probs = 0.25),
              q25_y = quantile(acc_y, probs = 0.25),
              q25_z = quantile(acc_z, probs = 0.25),
              q50_x = quantile(acc_x, probs = 0.50),
              q50_y = quantile(acc_y, probs = 0.50),
              q50_z = quantile(acc_z, probs = 0.50),
              q75_x = quantile(acc_x, probs = 0.75),
              q75_y = quantile(acc_y, probs = 0.75),
              q75_z = quantile(acc_z, probs = 0.75),
              cov_x_y = cov(acc_x, acc_y),
              cov_x_z = cov(acc_x, acc_z), 
              cov_y_z = cov(acc_y, acc_z),
              cor_x_y = cor(acc_x, acc_y),
              cor_x_z = cor(acc_x, acc_z),
              cor_y_z = cor(acc_y, acc_z),
              mean_diff_x_y = mean(acc_x-acc_y),
              mean_diff_x_z = mean(acc_x-acc_z),
              mean_diff_y_z = mean(acc_y-acc_z),
              sd_diff_x_y = sd(acc_x-acc_y),
              sd_diff_x_z = sd(acc_x-acc_z),
              sd_diff_y_z = sd(acc_y-acc_z),
              mean_amplitude_x = mean_amplitude(acc_x),
              mean_amplitude_y = mean_amplitude(acc_y),
              mean_amplitude_z = mean_amplitude(acc_z)) %>%
  ungroup()

```

### Prepare the dataset before the model training sequence

The classification model requires a "wide format" dataset, and the current dataset is in "long format". Here, we transform the dataset and add the statistical features.

```{r long_to_wide}
# Transform from long format to wide format
full_unobs_data <- unobs_raw_acc %>%
  add_column(idx = rep(1:bout_length, nrow(unobs_raw_acc)/100)) %>%
  dplyr::select(bout_id, idx, device_id, acc_x, acc_y, acc_z) %>%
  pivot_wider(names_from = idx, values_from = c(acc_x, acc_y, acc_z)) 

full_unobs_data <- left_join(full_unobs_data, 
                       unobs_raw_acc[, c("bout_id", "device_id", "UTC_datetime")],
                       by = c("device_id", "bout_id"))

full_unobs_data <- left_join(full_unobs_data, 
                       stat_feats,
                       by = c("device_id", "bout_id"))

```

## C - Behavioral classification of unobserved data

Here we use the previously built model to classify the behavior of individuals that were not observed.

::: callout-tip
#### ***Important!***

-   If you are classifying animal behavior **using your random forest model**, replace the file "gv_final_model_fit.rda" with your own random forest model.
:::

```{r upload_training_dataset}

################################################################
##### REPLACE THIS FILE WITH YOUR OWN RANDOM FOREST MODEL ######
################################################################

unobs_fit <- readRDS("../Data/gv_final_model_fit.rda")

# Separate - another chunk
full_unobs_data$start_int <- as.character(full_unobs_data$UTC_datetime)

predictions <-  predict(unobs_fit, full_unobs_data)

# Separate - another chunk
# Calculate a confidence score for each prediction
scores <- predict(unobs_fit, full_unobs_data, type='prob')

# What's still missing?!
# 1 - annotate this final part;
# 2 - add a part where the user can select if they want all the scores, or only the maximum score.


```
