---
title: "Carcasses and vulture feeding behavior"
author: "Kaija Gahm"
date: today
format:
  revealjs:
    slide-number: c
    embed-resources: true
editor: source
execute:
  echo: false
  message: false
  warning: false
---

```{r}
# SETUP; load and prep data
library(targets)
library(tidyverse)
library(here)
library(lubridate)
library(ggmap)
library(sf)
library(viridis)
library(mapview)
library(ggspatial)
library(readr)
source(here("R/functions.R"))
library(patchwork)
library(spatsoc)

tar_load(all_carcasses_annotated)
aca <- all_carcasses_annotated
tar_load(all_bouts_annotated)
aba <- all_bouts_annotated
tar_load(carcass_bouts)
tar_load(stations)
```

How many carcasses in each period, wild and inpa?
- Wild carcasses over time
How many bouts per carcass?
How many individuals per carcass?
Timeline of bouts at carcasses
- INPA only
- XXX need to go assign later wild feeding bouts back to the already-determined wild carcass locations

How accurate are the recorded carcass placement locations, based on the bouts? How far away are the bouts in general?
Does the distance of bouts from the recorded carcass location change over time?

## Carcass dynamics at feeding stations
How often are carcasses placed at feeding stations?
Are there days when more than one carcass is placed?
Overall visualization for carcasses placed over time
```{r}
aca <- aca %>%
  mutate(stationName = factor(stationName, levels = sort(unique(aca$stationName))))
glimpse(aca)

dg(stations) %>% dplyr::select(stationName, "stnlong" = long, "stnlat" = lat)
stn_carcass_timeline <- st_crop(aca, bbox_south) %>%
  filter(!cage, carcType == "inpa", stationName != "Other") %>%
  dplyr::select(datetime, dateOnly, year, stationName) %>%
  left_join(dg(stations) %>% 
              dplyr::select(stationName, "stnlong" = long, "stnlat" = lat)) %>%
  mutate(stationName = factor(stationName),
         stationName = fct_reorder(stationName, stnlat)) %>%
  ggplot(aes(x = datetime, y = stationName, col = stnlat))+
  scale_color_viridis_c()+
  geom_point(size = 3, alpha = 0.8)+
  facet_wrap(~year, scales = "free_x")+
  theme_bw()+
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        panel.spacing = unit(1, "lines"))+
  scale_x_datetime(date_breaks = "1 week",
                   date_labels = "%d %b")
#stn_carcass_timeline

mp <- st_crop(stations, bbox_south) %>%
  filter(stationName %in% (aca %>%
                             filter(!cage, carcType == "inpa", 
                                    stationName != "Other") %>%
                             pull(stationName))) %>%
  mutate(stationName = factor(stationName),
         stationName = fct_reorder(stationName, lat)) %>%
  ggplot()+
  annotation_map_tile(type = "cartolight", zoom = 10)+ 
  geom_sf(aes(col = lat), size = 3, alpha = 0.8)+
  scale_color_viridis_c()+
  ggrepel::geom_label_repel(
    aes(label = stationName,
        geometry = geometry),
    stat = "sf_coordinates",
    min.segment.length = 0.5,
    size = 3
  )+
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        axis.title.x = element_blank())+
  NULL
```
```{r}
stn_carcass_timeline | mp
```

```{r}
wild_carcass_timeline <- st_crop(aca, bbox_south) %>%
  st_transform("WGS84") %>%
  bind_cols(setNames(as.data.frame(st_coordinates(.)), c("long", "lat"))) %>%
  filter(carcType == "wild") %>%
  ggplot(aes(x = dateOnly, y = lat))+
  geom_point(size = 3, alpha = 0.8, aes(col = lat), pch = 1)+
  scale_color_viridis_c()+
  facet_wrap(~year, scales = "free_x")+
  theme_bw()+
  theme(axis.title.x = element_blank(),
        panel.spacing = unit(1, "lines"),
        legend.position = "none")+
  scale_x_date(date_breaks = "1 week",
               date_labels = "%d %b")+
  labs(y = "Latitude")

mp_wild <- st_crop(aca, bbox_south) %>%
  st_transform("WGS84") %>%
  bind_cols(setNames(as.data.frame(st_coordinates(.)), c("long", "lat"))) %>%
  filter(carcType == "wild") %>%
  ggplot()+
  annotation_map_tile(type = "cartolight", zoom = 10)+ 
  geom_sf(size = 3, alpha = 0.8, aes(col = lat), pch = 1)+
  scale_color_viridis_c()+
  theme(legend.position = "none")+
  NULL
```

```{r}
wild_carcass_timeline | mp_wild
```

```{r}
st_crop(aca, bbox_south) %>%
  st_transform("WGS84") %>%
  bind_cols(setNames(as.data.frame(st_coordinates(.)), c("long", "lat"))) %>%
  ggplot()+
  annotation_map_tile(type = "cartolight", zoom = 10)+ 
  geom_sf(size = 3, alpha = 0.8, aes(col = lat, shape = carcType))+
  scale_shape_manual(values = c(19, 1), name = "Carcass\ntype")+
  scale_color_viridis_c(guide = "none")+
  #theme(legend.position = "none")+
  facet_wrap(~year)+
  NULL
```

