---
title: "Carcasses and vulture feeding behavior"
author: "Kaija Gahm"
date: today
format:
  revealjs:
    slide-number: c
    embed-resources: true
editor: source
execute:
  echo: false
  message: false
  warning: false
---

```{r}
# SETUP; load and prep data
library(targets)
library(tidyverse)
library(here)
library(lubridate)
library(ggmap)
library(sf)
library(viridis)
library(mapview)
library(ggspatial)
library(readr)
source(here("R/functions.R"))
library(patchwork)
library(spatsoc)

tar_load(all_carcasses_annotated)
aca <- all_carcasses_annotated
tar_load(all_bouts_annotated)
aba <- all_bouts_annotated
tar_load(carcass_bouts)
tar_load(carcass_bouts_df)
tar_load(stations)
tar_load(bbox_south)
```

## How do vultures find carcasses?
- individual information
  - memory
  - searching
- social information 
  - in-flight (local enhancement)
  - following from the roost (information center hypothesis)
. . .

A combination of all of these!

## What factors affect reliance on different information sources?
H1: Resource predictability affects reliance on social information
. . .

Spatial and temporal predictability

  P1: Where resources are unpredictable, vultures will rely more on social information.
  P2: In areas with predictable resources, social information will not diminish in importance and will contribute



Predictability of carcasses in area
Size of carcass
Location (relative to roost sites; relative to overall pop. home range, etc.)
Dyadic interactions
Higher-order interactions
Signaling/circling



## Feeding station (INPA) carcasses

```{r}
aca <- aca %>%
  mutate(stationName = factor(stationName, levels = sort(unique(aca$stationName))))

dg(stations) %>% dplyr::select(stationName, "stnlong" = long, "stnlat" = lat)
stn_carcass_timeline <- st_crop(aca, bbox_south) %>%
  filter(!cage, carcType == "inpa", stationName != "Other") %>%
  dplyr::select(datetime, dateOnly, year, stationName) %>%
  left_join(dg(stations) %>% 
              dplyr::select(stationName, "stnlong" = long, "stnlat" = lat)) %>%
  mutate(stationName = factor(stationName),
         stationName = fct_reorder(stationName, stnlat)) %>%
  ggplot(aes(x = datetime, y = stationName, col = stnlat))+
  scale_color_viridis_c()+
  geom_point(size = 3, alpha = 0.8)+
  facet_wrap(~year, scales = "free_x")+
  theme_bw()+
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        panel.spacing = unit(1, "lines"))+
  scale_x_datetime(date_breaks = "1 week",
                   date_labels = "%d %b")
#stn_carcass_timeline

mp <- st_crop(stations, bbox_south) %>%
  filter(stationName %in% (aca %>%
                             filter(!cage, carcType == "inpa", 
                                    stationName != "Other") %>%
                             pull(stationName))) %>%
  mutate(stationName = factor(stationName),
         stationName = fct_reorder(stationName, lat)) %>%
  ggplot()+
  annotation_map_tile(type = "cartolight", zoom = 10)+ 
  geom_sf(aes(col = lat), size = 3, alpha = 0.8)+
  scale_color_viridis_c()+
  ggrepel::geom_label_repel(
    aes(label = stationName,
        geometry = geometry),
    stat = "sf_coordinates",
    min.segment.length = 0.5,
    size = 3
  )+
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        axis.title.x = element_blank())+
  NULL
```
```{r}
stn_carcass_timeline | mp
```

## Wild carcasses

```{r}
wild_carcass_timeline <- st_crop(aca, bbox_south) %>%
  st_transform("WGS84") %>%
  bind_cols(setNames(as.data.frame(st_coordinates(.)), c("long", "lat"))) %>%
  filter(carcType == "wild") %>%
  ggplot(aes(x = dateOnly, y = lat))+
  geom_point(size = 3, alpha = 0.8, aes(col = lat), pch = 1)+
  scale_color_viridis_c()+
  facet_wrap(~year, scales = "free_x")+
  theme_bw()+
  theme(axis.title.x = element_blank(),
        panel.spacing = unit(1, "lines"),
        legend.position = "none")+
  scale_x_date(date_breaks = "1 week",
               date_labels = "%d %b")+
  labs(y = "Latitude")

mp_wild <- st_crop(aca, bbox_south) %>%
  st_transform("WGS84") %>%
  bind_cols(setNames(as.data.frame(st_coordinates(.)), c("long", "lat"))) %>%
  filter(carcType == "wild") %>%
  ggplot()+
  annotation_map_tile(type = "cartolight", zoom = 10)+ 
  geom_sf(size = 3, alpha = 0.8, aes(col = lat), pch = 1)+
  scale_color_viridis_c()+
  theme(legend.position = "none")+
  NULL
```

```{r}
wild_carcass_timeline | mp_wild
```

## All carcasses

```{r}
st_crop(aca, bbox_south) %>%
  st_transform("WGS84") %>%
  bind_cols(setNames(as.data.frame(st_coordinates(.)), c("long", "lat"))) %>%
  ggplot()+
  annotation_map_tile(type = "cartolight", zoom = 10)+ 
  geom_sf(size = 3, alpha = 0.8, aes(col = lat, shape = carcType))+
  scale_shape_manual(values = c(19, 1), name = "Carcass\ntype")+
  scale_color_viridis_c(guide = "none")+
  #theme(legend.position = "none")+
  facet_wrap(~year)+
  NULL
```

## Carcass numbers
```{r}
aca %>%
  group_by(year, carcType) %>%
  summarize(n = n()) %>%
  ggplot(aes(x = factor(year), y = n, fill = carcType))+
  geom_col(position = position_dodge())+
  theme_bw()+
  scale_fill_manual(name = "Carcass\ntype", values = c("blue", "red"))+
  labs(y = "Number of carcasses",
       x = "Year")
```
## INPA carcass visits
```{r}
carcass_stats <- aca %>%
  filter(carcType == "inpa") %>%
  dplyr::select(carcID, X, Y, stationName, carcassWeight, "carcDatetime" = datetime, "carcDate" = dateOnly) %>%
  left_join(carcass_bouts_df, by = "carcID") %>%
  filter(!is.na(year)) %>%
  group_by(carcID, carcDate, carcDatetime, year) %>%
  summarize(nbouts = length(unique(boutID)),
            nindivs = length(unique(individualID)))

carcass_stats %>%
  ggplot(aes(x = nindivs, y = nbouts, col = factor(year)))+
  geom_point(size = 3, pch = 1)+
  geom_smooth(method = "lm")+
  theme_bw()+
  labs(y = "# recorded feeding bouts",
       x = "# individual vultures",
       caption = "(Within 72 hours and 750m of carcass placement at feeding station)")+
  scale_color_manual(name = "Year", values = c("skyblue", "skyblue4"))+
  theme(text = element_text(size = 16))
```

## INPA carcass visits

```{r}
carcass_stats %>%
  pivot_longer(cols = c("nindivs", "nbouts")) %>%
  mutate(name = case_when(name == "nbouts" ~ "Feeding bouts",
                          name == "nindivs" ~ "Individual vultures")) %>%
  ggplot(aes(x = factor(year), y = value, fill = factor(year)))+
  geom_boxplot()+
  scale_fill_manual(values = c("skyblue", "skyblue4"))+
  theme_bw()+
  facet_wrap(~name, scales = "free_y")+
  labs(x = "Year")+
  theme(axis.title.y = element_blank(),
        legend.position = "none",
        text = element_text(size = 16))
```



How many bouts per carcass?
How many individuals per carcass?
Timeline of bouts at carcasses
- INPA only
- XXX need to go assign later wild feeding bouts back to the already-determined wild carcass locations

How accurate are the recorded carcass placement locations, based on the bouts? How far away are the bouts in general?
Does the distance of bouts from the recorded carcass location change over time?



