---
title: "Carcasses and vulture feeding behavior"
author: "Kaija Gahm"
date: today
format:
  revealjs:
    slide-number: c
    embed-resources: true
editor: source
execute:
  echo: false
  message: false
  warning: false
---

```{r}
# SETUP
library(targets)
library(tidyverse)
library(here)
library(lubridate)
library(ggmap)
library(sf)
library(viridis)
library(mapview)
library(ggspatial)
library(readr)
source(here("R/functions.R"))
library(patchwork)
library(spatsoc)

carcasses_inpa <- readRDS(here("data/created/carcasses_inpa.RDS"))
stations <- readRDS(here("data/created/stations.RDS"))
feeding_bouts_station_2023 <- readRDS(here("data/ACC/2023_hf_period/created/feeding_bouts_station_2023.RDS"))
fbs_2023 <- feeding_bouts_station_2023[!map_lgl(feeding_bouts_station_2023, is.null)]

feeding_bouts_station_2024 <- readRDS(here("data/ACC/2024_hf_period/created/feeding_bouts_station_2024.RDS"))
fbs_2024 <- feeding_bouts_station_2024[!map_lgl(feeding_bouts_station_2024, is.null)]

bouts23 <- purrr::list_rbind(fbs_2023) %>%
  mutate(year = 2023,
         start = lubridate::ymd_hms(start),
         end = lubridate::ymd_hms(end),
         dateOnly = lubridate::ymd(dateOnly),
         tag_id = as.numeric(tag_id),
         individual_id = as.numeric(individual_id))
bouts24 <- fbs_2024 %>%
  map(., ~.x %>% mutate(tag_id = as.numeric(tag_id),
                        individual_id = as.numeric(individual_id))) %>%
  purrr::list_rbind() %>%
  mutate(year = 2024,
         start = lubridate::ymd_hms(start),
         end = lubridate::ymd_hms(end),
         dateOnly = lubridate::ymd(dateOnly))
bouts <- bind_rows(bouts23, bouts24) %>%
  sf::st_as_sf(crs = 32636)

write_rds(bouts, here("data/created/bouts.RDS"))
bouts <- readRDS(here("data/created/bouts.RDS"))

# Restrict to carcass placements during the time of year when we have hf-acc data
minmax_dates <- readRDS(here("data/created/minmax_dates.RDS"))

carcasses_23 <- carcasses_inpa %>% filter(datetime >= minmax_dates[1], datetime <= minmax_dates[2])
carcasses_24 <- carcasses_inpa %>% filter(datetime >= minmax_dates[3], datetime <= minmax_dates[4])
carcasses_focal <- bind_rows(carcasses_23, carcasses_24)

nrow(carcasses_focal) #104 feeding occurrences during the high-frequency ACC periods.

#For each carcass, let's look at feeding bouts within 750m and ~1 day of the carcass.
carcass_bouts <- map(1:nrow(carcasses_focal), ~{
  carcass <- carcasses_focal[.x,]
  id <- carcasses_focal$carcass_id[.x]
  distances <- as.numeric(st_distance(bouts, carcass))
  bouts <- bouts %>%
    mutate(carcass_id = id,
           dist_to_carcass = distances)
  keep_distance <- bouts %>%
    filter(dist_to_carcass <= 750)
  keep_time <- keep_distance %>%
    filter(start >= (carcass$datetime - hours(1)), # allow 1 hour before carcass
           end <= (carcass$datetime + hours(24))) %>% # allow 24 hours after carcass
    mutate(time_since_carcass = difftime(start, carcass$datetime, units = "hours"))
  return(keep_time)
}) # only one of these carcasses has a couple feeding bouts. We can look at this for now.
map_dbl(carcass_bouts, nrow) # varying numbers of feeding bouts for each carcass--many don't have any feeding bouts at all.

write_rds(carcass_bouts, here("data/created/carcass_bouts.RDS"))
carcass_bouts <- readRDS(here("data/created/carcass_bouts.RDS"))

# Create plots of feeding bouts relative to carcass locations, with feeding bouts colored by either distance or time

# Now let's see how the location of those feeding bouts corresponds to the location of the carcass as recorded by the INPA.

makeplots <- function(bouts, carc, type = "distance"){
  toplot <- bind_rows(carc %>%
                        mutate(type = "carcass"),
                      bouts %>%
                        mutate(type = "feeding_bout"))
  if(type == "time"){
    p <- ggplot(bouts)+
      geom_sf(aes(col = time_since_carcass),
              size = 5, alpha = 0.5, pch = 19)+
      scale_color_viridis(name = "Time since\nplacement (hr)", 
                          limits = c(-1, 25),
                          direction = -1)+
      theme_bw()+
      annotation_scale(location = "br")+
      geom_sf(data = carc, size = 5, pch = 8, col = "black")
  }else if(type == "distance"){
    p <- ggplot(bouts)+
      geom_sf(aes(col = dist_to_carcass),
              size = 5, alpha = 0.5, pch = 19)+
      scale_color_viridis(name = "Distance\nto carcass (m)", 
                          limits = c(0, 100),
                          option = "B",
                          direction = -1)+
      theme_bw()+
      annotation_scale(location = "br")+
      geom_sf(data = carc, size = 5, pch = 8, col = "black")
  }else(stop("`type` must be either 'time' or 'distance'"))
  return(p)
}

# Include the ones that have corresponding feeding bouts
toinclude <- which(map_dbl(carcass_bouts, nrow) > 0)
distanceplots <- map(toinclude, ~{
  makeplots(bouts = carcass_bouts[[.x]], carc = carcasses_focal[.x,], type = "distance")
})

timeplots <- map(toinclude, ~{
  makeplots(bouts = carcass_bouts[[.x]], carc = carcasses_focal[.x,], type = "time")
})
```

## Many carcasses don't have any recorded feeding bouts within 24 hours
```{r}
c_bouts <- purrr::list_rbind(carcass_bouts)

c_bouts %>%
  group_by(carcass_id) %>%
  summarize(bouts = n(),
            vultures = length(unique(individual_id)))
# XXX START HERE--need to formally merge carcass data with bout data by doing a left_join

carcasses_focal <- carcasses_focal %>%
  mutate(n_bouts = map_dbl(carcass_bouts, nrow))

carcasses_focal %>%
  ggplot(aes(x = n_bouts))+
  geom_histogram()+
  theme_classic()+
  labs(y = "Frequency",
       x = "Number of feeding bouts")
```

## Map of carcasses by number of feeding bouts
```{r}

```


## How accurate are the INPA's recorded carcass placement locations?

For carcasses placed during the high-frequency ACC periods, I identified feeding bouts based on ACC and compared the GPS location of those bouts to the recorded GPS location of the carcass.

## Some examples:
```{r}
distanceplots[[1]] + timeplots[[1]]
```

## Some examples:
```{r}
distanceplots[[4]] + timeplots[[4]]
```

## Some examples:
```{r}
distanceplots[[11]] + timeplots[[11]]
```

## How far away are the bouts in general?
```{r}
all_carcass_bouts <- purrr::list_rbind(carcass_bouts) %>%
  st_as_sf(crs = 32636)
all_carcass_bouts <- all_carcass_bouts %>%
  bind_cols(st_coordinates(all_carcass_bouts))
all_carcass_bouts %>%
  group_by(carcass_id) %>%
  mutate(mn = mean(dist_to_carcass)) %>%
  ggplot(aes(x = fct_reorder(factor(carcass_id), dist_to_carcass, mean), y = dist_to_carcass, fill = mn))+
  geom_boxplot(outlier.size = 0.5)+
  theme_bw()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())+
  scale_fill_viridis(name = "Mean distance\nto carcass", limits = c(0, 100))+
  labs(x = "Carcass",
       y = "Distance (m)",
       title = "Distance from feeding bouts to recorded carcass location",
       sutbtitle = "For feeding bouts within 500m",
       caption = "Includes carcasses that had at least 1 feeding bout within 750m\n Includes feeding bouts from 1 hour before through 24 hours after carcass placement.")
```

## Does the distance of bouts from the recorded carcass location change over time?
```{r}
#| warning: false
all_carcass_bouts %>%
  group_by(carcass_id) %>%
  filter(n() > 4) %>%
  ggplot(aes(x = time_since_carcass, y = dist_to_carcass, group = carcass_id))+
  theme_bw()+
  geom_smooth(method = "lm")+
  geom_point(alpha = 0.5, pch = 1)+
  facet_wrap(~carcass_id, scales = "free_x")+
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())+
  labs(y = "Distance (m)",
       x = "Time since carcass placement (hours)",
       caption = "Includes carcasses that had at least 5 feeding bouts within 750m.\n Includes feeding bouts from 1 hour before through 24 hours after carcass placement.")
```

## Looks like the carcass locations are generally pretty accurate, and the carcass doesn't necessarily get dragged farther away over time!
A 100m distance threshold seems like a decent buffer for the INPA-provided locations.

## Let's understand more about the dynamics of carcasses at feeding stations.

- How often are carcasses placed at feeding stations?
- How regular is the carcass placement?
- How much does carcass cadence vary between feeding stations?

## Strangely, the INPA carcass data includes some points in the sea

```{r}
bbox <- st_bbox(carc_inpa)

ggplot(carc_inpa) +
  annotation_map_tile(zoom = 7, type = "cartolight") +
  geom_sf(show.legend = "point", inherit.aes = FALSE, alpha = 0.2, size = 2,
          col = "blue")+
  ggtitle("Carcass locations, 2018-2024")
```

## After removing those points:

```{r}
carc_inpa_edited <- carc_inpa %>%
  filter(!(long < 34.5 & lat > 33.0))

bbox <- st_bbox(carc_inpa_edited)

ggplot(carc_inpa_edited) +
  annotation_map_tile(zoom = 7, type = "cartolight") +
  geom_sf(show.legend = "point", inherit.aes = FALSE, alpha = 0.2, size = 2, 
          col = "blue")+
  ggtitle("Carcass locations, 2018-2024")
```

## The carcasses in the sea are all from two dates on August 2024. Typo? GPS scrambling?
```{r}
carc_inpa %>%
  filter(long < 34.5 & lat > 33.0) # Most are Carmel, one is Tzaror
```

## Feeding bouts, both at and outside of feeding stations
```{r}
bouts <- bouts %>%
  bind_cols(st_coordinates(.)) # here is all our feeding bout information
table(bouts$station) # includes bouts at and away from feeding stations
# XXX need to go back to the targets pipeline and change how we define station locations.
#head(feeding_bouts_station)
```

```{r}
# All carcasses
#mapview(feeding_bouts_station, zcol = "station") # we have lots of feeding information for a really wide area!

feeding_cropped <- st_crop(feeding_bouts_station, st_bbox(carc_inpa_edited))

# Let's make a plot requesting tiles just for the bbox
ggplot(feeding_cropped)+
  annotation_map_tile(zoom = 9, type = "hotstyle") +
  geom_sf(aes(col = station), alpha = 0.9, pch = 1)+
  scale_color_manual(name = "Station", values = c("red", "blue"))+
  facet_wrap(~year)+
  coord_sf(datum = st_crs(32636))
```

## Zooming in on the south
```{r}
new_bbox <- st_bbox(feeding_cropped)
new_bbox[4] <- 3500000
new_bbox[2] <- 3350000

feeding_cropped_south <- st_crop(feeding_cropped, new_bbox)

#rosm::osm.types()
ggplot(feeding_cropped_south)+
  annotation_map_tile(zoom = 9, type = "hotstyle") +
  geom_sf(aes(col = station), alpha = 0.9, pch = 1)+
  scale_color_manual(name = "Station", values = c("red", "blue"))+
  facet_wrap(~year)+
  coord_sf(datum = st_crs(32636))
```
## Can we use spatsoc to group the feeding bouts together into non-station carcass locations?
```{r}
ns_bouts <- feeding_cropped_south %>%
  filter(station == FALSE) %>%
  arrange(timestamp)
dim(ns_bouts)
st_crs(ns_bouts)
st_distance(ns_bouts[1,],  ns_bouts[2,]) # confirms that the units are in meters, 27,431 meters = ~27.4 km, which looks about right based on the map:
mapview(ns_bouts[1:2,])

# Prepare data for analysis with spatsoc
ns_bouts$timestamp <- as.POSIXct(ns_bouts$timestamp)
ns_bouts <- data.table::data.table(ns_bouts)
class(ns_bouts$timestamp)

group_times(ns_bouts, datetime = 'timestamp', threshold = '24 hours') # group into 24 hour groups
group_pts(ns_bouts, threshold = 100, id ='device_id', coords = c('X', 'Y'), timegroup = 'timegroup')
length(unique(ns_bouts$group)) # yessssss

length(unique(ns_bouts$timegroup)) 
length(unique(ns_bouts$dateOnly)) # okay, we only have 67 days of data, so it makes sense that this is how many groups we ended up with.
length(unique(ns_bouts$group)) # way more groups than days--this is as expected, because there will be multiple carcasses per day!
# XXX how many non-station carcasses per day?

ns_bouts %>%
  group_by(dateOnly, timegroup) %>%
  summarize(n_ns_carcasses = length(unique(group))) %>%
  arrange(desc(n_ns_carcasses)) # does this seem reasonable?? Gut check.
# XXX will have to deal with multi-day persistent carcasses

# convert back to sf object for mapping
ns_bouts <- as.data.frame(ns_bouts) %>%
  st_as_sf(crs = 32636)
```

```{r}
ns_carcasses <- ns_bouts %>%
  group_by(year, group) %>%
  summarize(geometry = st_union(geometry),
            dateOnly = dateOnly[1]) %>%
  st_centroid() %>%
  ungroup() %>%
  bind_cols(st_coordinates(.))
dim(ns_bouts)
dim(ns_carcasses)
head(ns_carcasses)
```

```{r}
ggplot()+
  annotation_map_tile(zoom = 9, type = "hotstyle") +
  geom_sf(data = ns_bouts, alpha = 0.5, pch = 1, col = "black")+
  geom_sf(data = ns_carcasses, alpha = 0.5, pch = 4, col = "red")+
  facet_wrap(~year)+
  labs(title = "Non-station carcass locations (inferred)",
       caption = "Black points represent feeding bouts. Red x's represent the centroids\nof groups of feeding bouts (24 hours, 100m clusters), i.e. putative\ncarcass locations.")+
  theme(plot.caption = element_text(hjust = 0))
```

## With the at-station carcasses added

```{r}
s_carcasses_match <- carc_inpa %>%
  bind_cols(st_coordinates(.)) %>%
  filter(Y >= min(ns_carcasses$Y), Y <= max(ns_carcasses$Y))

ggplot()+
  annotation_map_tile(zoom = 9, type = "hotstyle") +
  geom_sf(data = ns_bouts, alpha = 0.5, pch = 1, col = "black")+
  geom_sf(data = ns_carcasses, alpha = 0.5, pch = 4, col = "red")+
  geom_sf(data = s_carcasses_match, alpha = 0.3, pch = 4, col = "blue")+
  facet_wrap(~year) # XXX some of these are suspiciously close to the feeding station carcasses. I suspect I need to revise my definition of non-station. I have coordinates for the feeding stations in the south, so I can use that maybe? But then what if carcasses were sometimes placed outside of feeding stations?
```


## How often are carcasses placed at feeding stations?

Coordinates of feeding stations in the south: XXX READ IN

Coordinates of feeding stations in the north: XXX STILL NEED THIS FILE

Until we have a full list of feeding station coordinates, let's use the real carcass data to infer the position of feeding stations.

Let's first see where carcasses were recorded as being placed.

```{r}
#| echo: false
#| message: false
library(tidyverse)
library(sf)
library(mapview)
library(here)
library(readxl)
```

```{r}
#| echo: false
#| warning: false
#| message: false
feed <- read_excel(here("data/FeedingData from 2018_2024_Translated.xlsx")) %>% 
  select(contains("RTG"), contains("WGS84"), "date" = `Date Event`, "time" = `Event time`) %>%
  rename("inspectionArea" = `Inspection area - RTG`,
         "merhavRTG" = `Merhav - RTG`,
         "district" = `RTG district`,
         "long" = `WGS84 - LONG`,
         "lat" = `WGS84 - LAT`) %>%
  mutate(datetime = lubridate::ymd_hms(paste0(as.character(date), " ", substr(as.character(time), 12, 19)))) %>%
  mutate(year = lubridate::year(datetime))

feed_sf <- feed %>%
  st_as_sf(coords = c("long", "lat"), crs = "WGS84") %>%
  st_transform(32636)

ggplot(feed_sf)+
  geom_sf(alpha = 0.2, size = 2, col = "blue")+
  theme_minimal()+
  ggtitle("Carcass locations, 2018-2024")
```

We notice that there are definitely a lot of repeat placements. As a quick and dirty way of quantifying feeding stations, let's cluster them together.

Buffering by 300m (arbitrary) and then clustering and intersecting and coloring by cluster gives us this:

```{r}
#| echo: false
carc_inpa_buffered <- st_buffer(feed_sf, 300)
parts <- st_cast(st_union(carc_inpa_buffered),"POLYGON")
clust <- unlist(st_intersects(carc_inpa_buffered, parts))
diss <- cbind(carc_inpa_buffered, clust) %>%
  group_by(clust) %>%
  st_transform("WGS84")

#mapview(diss, zcol = "clust")
```

Examine the carcass placement over time at these "stations"
```{r}
#| echo: false
#| message: false
ggplot(diss, aes(x = datetime, y = clust, col = factor(clust)))+
  geom_point()+
  theme_bw()+
  ylab("Feeding Station")+
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())+
  labs(title = "Carcass placement at ~feeding stations",
       caption = "Feeding station locations defined by buffering and merging nearby coordinates.")+
  scale_x_datetime(date_breaks = "6 months",
                   date_labels = "%m/%y")
```

...and by year:
```{r}
#| echo: false
#| message: false
diss %>%
  group_by(year, clust) %>%
  filter(n() > 4) %>%
  ggplot(aes(x = datetime, y = factor(clust), 
             col = factor(clust)))+
  geom_point()+
  facet_wrap(~year, scales = "free") +
  theme_bw()+
  ylab("Feeding Station")+
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())+
  labs(title = "Carcass placement at ~feeding stations, by year",
       caption = "Feeding stations created by buffering and merging nearby coordinates.")+
  scale_x_datetime(date_breaks = "2 months",
                   date_labels = "%b")
```

Let's take a look at the intervals between carcass depositions
```{r}
#| echo: false
#| message: false
stats <- diss %>%
  arrange(clust, datetime) %>%
  group_by(clust) %>%
  mutate(interval = difftime(datetime, lag(datetime), units = "days"))

stats %>%
  group_by(clust) %>%
  filter(n() > 4) %>%
  filter(!is.na(interval)) %>%
  ggplot(aes(x = fct_reorder(factor(clust), interval, .fun = median), 
             y = interval))+
  geom_boxplot(outlier.shape = NA)+
  theme_minimal()+
  geom_jitter(alpha = 0.3, aes(col = factor(clust)), width = 0.2)+
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())+
  ylab("Interval (days)")+
  xlab("Feeding station")
```

Top 5 by year:
```{r}
#| echo: false
#| message: false
# Now picking the top 5
stats %>%
  filter(!is.na(interval)) %>%
  group_by(clust) %>%
  mutate(size = n()) %>%
  ungroup() %>%
  filter(size >= sort(unique(size), TRUE)[5])  %>%
  arrange(desc(size)) %>%
  ggplot(aes(x = fct_reorder(factor(clust), interval, .fun = median), 
             y = interval))+
  geom_boxplot(outlier.shape = NA)+
  theme_minimal()+
  geom_jitter(alpha = 0.3, aes(col = factor(clust)), width = 0.2)+
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())+
  ylab("Interval (days)")+
  xlab("Feeding station")+
  facet_wrap(~year, scales = "free_y")+
  ggtitle("Top 5 most frequent overall feeding stations")
```

```{r}
#| echo: false
#| message: false
lines <- stats %>%
  st_drop_geometry() %>%
  filter(!is.na(interval)) %>%
  group_by(clust) %>%
  mutate(size = n()) %>%
  ungroup() %>%
  filter(size >= sort(unique(size), TRUE)[5]) %>%
  group_by(clust, year) %>%
  summarize(mean = mean(interval))

stats %>%
  filter(!is.na(interval)) %>%
  group_by(clust) %>%
  mutate(size = n()) %>%
  ungroup() %>%
  filter(size >= sort(unique(size), TRUE)[5])  %>%
  arrange(desc(size)) %>%
  ggplot(aes(x = interval, group = clust, col = factor(clust)))+
  geom_density(linewidth = 1)+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.text.y = element_blank())+
  xlab("Interval (days)")+
  facet_wrap(~year, scales = "free")+
  geom_vline(data = lines, aes(xintercept = mean, col = factor(clust)), linetype = 2, linewidth = 0.5)+
  labs(title = "Frequency at 5 overall most provisioned feeding stations, by year")+
  ylab("")
```

So, depending on the feeding station, carcasses are typically deposited every 5-100 days. At the top/most frequent feeding stations, carcasses are placed about every 3-40 days.

How often do we see carcasses being placed less than 5 days apart?
```{r}
threshes <- seq(from = 1, to = 5, by = 1)
dat <- purrr::map(threshes, ~stats %>% filter(interval <= .x) %>% mutate(thresh = .x)) %>% purrr::list_rbind()

dat %>%
  st_drop_geometry() %>%
  group_by(clust, thresh) %>%
  add_count() %>%
  ggplot(aes(x = thresh, y = n, group = factor(clust), col = factor(clust)))+
  geom_line()+
  scale_x_continuous(breaks = threshes)+
  theme_minimal()+
  ylab("Number of carcass placements")+
  xlab("<= n days apart")
```

So, while most carcasses are provisioned less often than every 5 days, there are some sites where food is often available every day!

## Are there days when more than one carcass is placed?
```{r}
head(stats)
carcasses_per_day <- stats %>%
  st_drop_geometry() %>%
  group_by(year, date) %>%
  summarize(carc_per_day = n())
carcasses_per_day %>%
  ggplot(aes(x = factor(carc_per_day)))+
  geom_bar()+
  theme_minimal()+
  ylab("Days")+
  xlab("Carcasses per day")+
  facet_wrap(~year)+
  labs(caption = "Sometimes, more than one carcass\nis deposited per day\n(anywhere in the country)")
```

## Overall visualization of carcasses placed over time
```{r}
over_time <- stats %>%
  st_drop_geometry() %>%
  mutate(month = lubridate::month(datetime),
         day = lubridate::day(datetime)) %>%
  group_by(month, day, year, date) %>%
  summarize(n = n())

over_time %>%
  ggplot(aes(y = n, x = lubridate::ymd(date)))+
  geom_line(alpha = 0.2)+
  geom_point(aes(col = n))+
  scale_color_viridis_c(name = "")+
  theme_minimal()+
  facet_wrap(~year, scales = "free_x", ncol = 2)+
  scale_x_date(date_breaks = "3 months",
               date_labels = "%b")+
  ggtitle("Carcasses over time")+
  ylab("Number of carcasses")+
  xlab("Date")
```

## Summary stats for how long carcasses last
We can do this either by looking at the vultures that are nearby the stated location, or we can look at feeding bouts nearby. It would be better to look at feeding bouts in order to quantify duration, I think, so I'm going to wait on this until I have the data processed.
