---
title: "Report"
author: "Kaija Gahm"
format: html
editor: source
---

## How often are carcasses placed at feeding stations?

Coordinates of feeding stations in the south: XXX READ IN

Coordinates of feeding stations in the north: XXX STILL NEED THIS FILE

Until we have a full list of feeding station coordinates, let's use the real carcass data to infer the position of feeding stations.

Let's first see where carcasses were recorded as being placed.

```{r}
#| echo: false
#| message: false
library(tidyverse)
library(sf)
library(mapview)
library(here)
library(readxl)
```

```{r}
#| echo: false
#| warning: false
#| message: false
feed <- read_excel(here("data/FeedingData from 2018_2024_Translated.xlsx")) %>% 
  select(contains("RTG"), contains("WGS84"), "date" = `Date Event`, "time" = `Event time`) %>%
  rename("inspectionArea" = `Inspection area - RTG`,
         "merhavRTG" = `Merhav - RTG`,
         "district" = `RTG district`,
         "long" = `WGS84 - LONG`,
         "lat" = `WGS84 - LAT`) %>%
  mutate(datetime = lubridate::ymd_hms(paste0(as.character(date), " ", substr(as.character(time), 12, 19)))) %>%
  mutate(year = lubridate::year(datetime))

feed_sf <- feed %>%
  st_as_sf(coords = c("long", "lat"), crs = "WGS84") %>%
  st_transform(32636)

ggplot(feed_sf)+
  geom_sf(alpha = 0.2, size = 2, col = "blue")+
  theme_minimal()+
  ggtitle("Carcass locations, 2018-2024")
```

We notice that there are definitely a lot of repeat placements. As a quick and dirty way of quantifying feeding stations, let's cluster them together.

Buffering by 300m (arbitrary) and then clustering and intersecting and coloring by cluster gives us this:

```{r}
#| echo: false
fs_buffered <- st_buffer(feed_sf, 300)
parts <- st_cast(st_union(fs_buffered),"POLYGON")
clust <- unlist(st_intersects(fs_buffered, parts))
diss <- cbind(fs_buffered, clust) %>%
  group_by(clust) %>%
  st_transform("WGS84")

#mapview(diss, zcol = "clust")
```

Examine the carcass placement over time at these "stations"
```{r}
#| echo: false
#| message: false
ggplot(diss, aes(x = datetime, y = clust, col = factor(clust)))+
  geom_point()+
  theme_bw()+
  ylab("Feeding Station")+
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())+
  labs(title = "Carcass placement at ~feeding stations",
       caption = "Feeding stations created by buffering and merging nearby coordinates.")+
  scale_x_datetime(date_breaks = "6 months",
                   date_labels = "%m/%y")
```

...and by year:
```{r}
#| echo: false
#| message: false
diss %>%
  group_by(year, clust) %>%
  filter(n() > 4) %>%
  ggplot(aes(x = datetime, y = factor(clust), 
             col = factor(clust)))+
  geom_point()+
  facet_wrap(~year, scales = "free") +
  theme_bw()+
  ylab("Feeding Station")+
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())+
  labs(title = "Carcass placement at ~feeding stations, by year",
       caption = "Feeding stations created by buffering and merging nearby coordinates.")+
  scale_x_datetime(date_breaks = "2 months",
                   date_labels = "%b")
```

Let's take a look at the intervals between carcass depositions
```{r}
#| echo: false
#| message: false
stats <- diss %>%
  arrange(clust, datetime) %>%
  group_by(clust) %>%
  mutate(interval = difftime(datetime, lag(datetime), units = "days"))

stats %>%
  group_by(clust) %>%
  filter(n() > 4) %>%
  filter(!is.na(interval)) %>%
  ggplot(aes(x = fct_reorder(factor(clust), interval, .fun = median), 
             y = interval))+
  geom_boxplot(outlier.shape = NA)+
  theme_minimal()+
  geom_jitter(alpha = 0.3, aes(col = factor(clust)), width = 0.2)+
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())+
  ylab("Interval (days)")+
  xlab("Feeding station")
```

Top 5 by year:
```{r}
#| echo: false
#| message: false
# Now picking the top 5
stats %>%
  filter(!is.na(interval)) %>%
  group_by(clust) %>%
  mutate(size = n()) %>%
  ungroup() %>%
  filter(size >= sort(unique(size), TRUE)[5])  %>%
  arrange(desc(size)) %>%
  ggplot(aes(x = fct_reorder(factor(clust), interval, .fun = median), 
             y = interval))+
  geom_boxplot(outlier.shape = NA)+
  theme_minimal()+
  geom_jitter(alpha = 0.3, aes(col = factor(clust)), width = 0.2)+
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())+
  ylab("Interval (days)")+
  xlab("Feeding station")+
  facet_wrap(~year, scales = "free_y")+
  ggtitle("Top 5 most frequent overall feeding stations")
```

```{r}
#| echo: false
#| message: false
lines <- stats %>%
  st_drop_geometry() %>%
  filter(!is.na(interval)) %>%
  group_by(clust) %>%
  mutate(size = n()) %>%
  ungroup() %>%
  filter(size >= sort(unique(size), TRUE)[5]) %>%
  group_by(clust, year) %>%
  summarize(mean = mean(interval))

stats %>%
  filter(!is.na(interval)) %>%
  group_by(clust) %>%
  mutate(size = n()) %>%
  ungroup() %>%
  filter(size >= sort(unique(size), TRUE)[5])  %>%
  arrange(desc(size)) %>%
  ggplot(aes(x = interval, group = clust, col = factor(clust)))+
  geom_density(linewidth = 1)+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.text.y = element_blank())+
  xlab("Interval (days)")+
  facet_wrap(~year, scales = "free")+
  geom_vline(data = lines, aes(xintercept = mean, col = factor(clust)), linetype = 2, linewidth = 0.5)+
  labs(title = "Frequency at 5 overall most provisioned feeding stations, by year")+
  ylab("")
```

So, depending on the feeding station, carcasses are typically deposited every 5-100 days. At the top/most frequent feeding stations, carcasses are placed about every 3-40 days.

How often do we see carcasses being placed less than 5 days apart?
```{r}
threshes <- seq(from = 1, to = 5, by = 1)
dat <- purrr::map(threshes, ~stats %>% filter(interval <= .x) %>% mutate(thresh = .x)) %>% purrr::list_rbind()

dat %>%
  st_drop_geometry() %>%
  group_by(clust, thresh) %>%
  add_count() %>%
  ggplot(aes(x = thresh, y = n, group = factor(clust), col = factor(clust)))+
  geom_line()+
  scale_x_continuous(breaks = threshes)+
  theme_minimal()+
  ylab("Number of carcass placements")+
  xlab("<= n days apart")
```

So, while most carcasses are provisioned less often than every 5 days, there are some sites where food is often available every day!

## Are there days when more than one carcass is placed?
```{r}
head(stats)
carcasses_per_day <- stats %>%
  st_drop_geometry() %>%
  group_by(year, date) %>%
  summarize(carc_per_day = n())
carcasses_per_day %>%
  ggplot(aes(x = factor(carc_per_day)))+
  geom_bar()+
  theme_minimal()+
  ylab("Days")+
  xlab("Carcasses per day")+
  facet_wrap(~year)+
  labs(caption = "Sometimes, more than one carcass\nis deposited per day\n(anywhere in the country)")
```

## Overall visualization of carcasses placed over time
```{r}
over_time <- stats %>%
  st_drop_geometry() %>%
  mutate(month = lubridate::month(datetime),
         day = lubridate::day(datetime)) %>%
  group_by(month, day, year, date) %>%
  summarize(n = n())

over_time %>%
  ggplot(aes(y = n, x = lubridate::ymd(date)))+
  geom_line(alpha = 0.2)+
  geom_point(aes(col = n))+
  scale_color_viridis_c(name = "")+
  theme_minimal()+
  facet_wrap(~year, scales = "free_x", ncol = 2)+
  scale_x_date(date_breaks = "3 months",
               date_labels = "%b")+
  ggtitle("Carcasses over time")+
  ylab("Number of carcasses")+
  xlab("Date")
```

## Summary stats for how long carcasses last
We can do this either by looking at the vultures that are nearby the stated location, or we can look at feeding bouts nearby. It would be better to look at feeding bouts in order to quantify duration, I think, so I'm going to wait on this until I have the data processed.
