---
title: "Carcasses and vulture feeding behavior"
author: "Kaija Gahm"
date: today
format:
  revealjs:
    slide-number: c
editor: source
execute:
  echo: false
  message: false
  warning: false
---

```{r}
# SETUP
library(targets)
library(tidyverse)
library(here)
library(lubridate)
library(ggmap)
library(sf)
library(viridis)
library(mapview)
library(ggspatial)
library(readr)
source(here("R/functions.R"))
library(patchwork)
library(spatsoc)

tar_load(fs)
# tar_load(feeding_bouts_station)
# bouts24 <- feeding_bouts_station
# bouts23 <- readRDS(here("data/ACC/2023_hf_period/created/feeding_bouts_station_2023.RDS"))
# 
# bouts <- purrr::list_rbind(bouts24) %>%
#   mutate(year = 2024,
#          start = lubridate::ymd_hms(start),
#          end = lubridate::ymd_hms(end),
#          dateOnly = lubridate::ymd(dateOnly)) %>%
#   bind_rows(purrr::list_rbind(bouts23) %>%
#               mutate(year = 2023,
#                      tag_id = as.numeric(tag_id),
#                      dateOnly = lubridate::ymd(dateOnly),
#                      individual_id = as.numeric(individual_id))) %>%
#   sf::st_as_sf(crs = "WGS84") %>%
#   sf::st_transform(32636)
# write_rds(bouts, here("data/bouts.RDS"))
bouts <- readRDS(here("data/bouts.RDS"))

# Restrict to carcass placements during the time of year
mindate_23 <- bouts %>% filter(year == 2023) %>% pull(start) %>% min()
maxdate_23 <- bouts %>% filter(year == 2023) %>% pull(start) %>% max()

mindate_24 <- bouts %>% filter(year == 2024) %>% pull(start) %>% min()
maxdate_24 <- bouts %>% filter(year == 2024) %>% pull(start) %>% max()

fs_focal_23 <- fs %>% filter(datetime >= mindate_23, datetime <= maxdate_23)
fs_focal_24 <- fs %>% filter(datetime >= mindate_24, datetime <= maxdate_24)
fs_focal <- bind_rows(fs_focal_23, fs_focal_24)

# nrow(fs_focal) #104 feeding occurrences during the high-frequency ACC periods.

# For each carcass, let's look at feeding bouts within 750m and ~1 day of the carcass.
# carcass_bouts <- map(1:nrow(fs_focal), ~{
#   carcass <- fs_focal[.x,]
#   id <- fs_focal$carcass_id[.x]
#   distances <- as.numeric(st_distance(bouts, carcass))
#   bouts <- bouts %>%
#     mutate(carcass_id = id,
#            dist_to_carcass = distances)
#   keep_distance <- bouts %>%
#     filter(dist_to_carcass <= 750)
#   keep_time <- keep_distance %>%
#     filter(start >= (carcass$datetime - hours(1)), # allow 1 hour before carcass
#            end <= (carcass$datetime + hours(24))) %>% # allow 24 hours after carcass
#     mutate(time_since_carcass = difftime(start, carcass$datetime, units = "hours"))
#   return(keep_time)
# }) # only one of these carcasses has a couple feeding bouts. We can look at this for now.
# write_rds(carcass_bouts, here("data/carcass_bouts.RDS"))
carcass_bouts <- readRDS(here("data/carcass_bouts.RDS"))

# Create plots of feeding bouts relative to carcass locations, with feeding bouts colored by either distance or time
toinclude <- which(map_dbl(carcass_bouts, nrow) > 0)

# Now let's see how the location of those feeding bouts corresponds to the location of the carcass as recorded by the INPA.

makeplots <- function(bouts, carc, type = "distance"){
  toplot <- bind_rows(carc %>%
                        mutate(type = "carcass"),
                      bouts %>%
                        mutate(type = "feeding_bout"))
  if(type == "time"){
    p <- ggplot(bouts)+
      geom_sf(aes(col = time_since_carcass),
              size = 5, alpha = 0.5, pch = 19)+
      scale_color_viridis(name = "Time since\nplacement (hr)", 
                          limits = c(-1, 25),
                          direction = -1)+
      theme_bw()+
      annotation_scale(location = "br")+
      geom_sf(data = carc, size = 5, pch = 8, col = "black")
  }else if(type == "distance"){
    p <- ggplot(bouts)+
      geom_sf(aes(col = dist_to_carcass),
              size = 5, alpha = 0.5, pch = 19)+
      scale_color_viridis(name = "Distance\nto carcass (m)", 
                          limits = c(0, 100),
                          option = "B",
                          direction = -1)+
      theme_bw()+
      annotation_scale(location = "br")+
      geom_sf(data = carc, size = 5, pch = 8, col = "black")
  }else(stop("`type` must be either 'time' or 'distance'"))
  return(p)
}
distanceplots <- map(toinclude, ~{
  makeplots(bouts = carcass_bouts[[.x]], carc = fs_focal[.x,], type = "distance")
})

timeplots <- map(toinclude, ~{
  makeplots(bouts = carcass_bouts[[.x]], carc = fs_focal[.x,], type = "time")
})
```

## How accurate are the INPA's recorded carcass placement locations?

For carcasses placed during the high-frequency ACC periods, I identified feeding bouts based on ACC and compared the GPS location of those bouts to the recorded GPS location of the carcass.

## Some examples:
```{r}
distanceplots[[1]] + timeplots[[1]]
```

## Some examples:
```{r}
distanceplots[[4]] + timeplots[[4]]
```

## Some examples:
```{r}
distanceplots[[11]] + timeplots[[11]]
```

## How far away are the bouts in general?
```{r}
all_carcass_bouts <- purrr::list_rbind(carcass_bouts)
all_carcass_bouts %>%
  group_by(carcass_id) %>%
  mutate(mn = mean(dist_to_carcass)) %>%
  ggplot(aes(x = fct_reorder(factor(carcass_id), dist_to_carcass, mean), y = dist_to_carcass, fill = mn))+
  geom_boxplot(outlier.size = 0.5)+
  theme_bw()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())+
  scale_fill_viridis(name = "Mean distance\nto carcass", limits = c(0, 100))+
  labs(x = "Carcass",
       y = "Distance (m)",
       title = "Distance from feeding bouts to recorded carcass location",
       sutbtitle = "For feeding bouts within 500m",
       caption = "Includes carcasses that had at least 1 feeding bout within 750m\n Includes feeding bouts from 1 hour before through 24 hours after carcass placement.")
```

## Does the distance of bouts from the recorded carcass location change over time?
```{r}
#| warning: false
all_carcass_bouts %>%
  group_by(carcass_id) %>%
  filter(n() > 4) %>%
  ggplot(aes(x = time_since_carcass, y = dist_to_carcass, group = carcass_id))+
  theme_bw()+
  geom_smooth(method = "lm")+
  geom_point(alpha = 0.5, pch = 1)+
  facet_wrap(~carcass_id, scales = "free_x")+
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())+
  labs(y = "Distance (m)",
       x = "Time since carcass placement (hours)",
       caption = "Includes carcasses that had at least 5 feeding bouts within 750m.\n Includes feeding bouts from 1 hour before through 24 hours after carcass placement.")
```

## Looks like the carcass locations are generally pretty accurate, and the carcass doesn't necessarily get dragged farther away over time!
A 100m distance threshold seems like a decent buffer for the INPA-provided locations.

## Let's understand more about the dynamics of carcasses at feeding stations.

- How often are carcasses placed at feeding stations?
- How regular is the carcass placement?
- How much does carcass cadence vary between feeding stations?

## Strangely, the INPA carcass data includes some points in the sea

```{r}
bbox <- st_bbox(fs)

ggplot(fs) +
  annotation_map_tile(zoom = 7, type = "cartolight") +
  geom_sf(show.legend = "point", inherit.aes = FALSE, alpha = 0.2, size = 2,
          col = "blue")+
  ggtitle("Carcass locations, 2018-2024")
```

## After removing those points:

```{r}
fs_edited <- fs %>%
  filter(!(long < 34.5 & lat > 33.0))

bbox <- st_bbox(fs_edited)

ggplot(fs_edited) +
  annotation_map_tile(zoom = 7, type = "cartolight") +
  geom_sf(show.legend = "point", inherit.aes = FALSE, alpha = 0.2, size = 2, 
          col = "blue")+
  ggtitle("Carcass locations, 2018-2024")
```

## The carcasses in the sea are all from two dates on August 2024. Typo?
```{r}
fs %>%
  filter(long < 34.5 & lat > 33.0)
```
## Feeding bouts, both at and outside of feeding stations
```{r}
targets::tar_load(feeding_bouts_station) # these were already assigned station vs. not using a 100m threshold, so I'm glad that the above analysis confirms that is a reasonable threshold!
feeding_bouts_station_2024 <- purrr::list_rbind(feeding_bouts_station) %>%
  mutate(start = lubridate::ymd_hms(start),
         end = lubridate::ymd_hms(end),
         timestamp = lubridate::ymd_hms(timestamp),
         dateOnly = lubridate::ymd(dateOnly))
feeding_bouts_station_2023 <- readRDS(here("data/ACC/2023_hf_period/created/feeding_bouts_station_2023.RDS")) %>%
  purrr::list_rbind() %>% mutate(tag_id = as.numeric(tag_id),
                                 individual_id = as.numeric(individual_id),
                                 dateOnly = lubridate::ymd(dateOnly))

feeding_bouts_station <- bind_rows(feeding_bouts_station_2023, feeding_bouts_station_2024) %>% mutate(year = lubridate::year(timestamp)) %>%
  select(bout_id, device_id, pred, .pred_Eating, start, end, tag_id, timestamp, dateOnly, ground_speed, individual_id, geometry, station, year) %>%
  st_as_sf(crs = "WGS84")
#head(feeding_bouts_station)
```

```{r}
# All carcasses
#mapview(feeding_bouts_station, zcol = "station") # we have lots of feeding information for a really wide area!

feeding_cropped <- st_crop(feeding_bouts_station, st_bbox(st_transform(fs_edited, "WGS84")))

# Let's make a plot requesting tiles just for the bbox
ggplot(feeding_cropped)+
  annotation_map_tile(zoom = 9, type = "hotstyle") +
  geom_sf(aes(col = station), alpha = 0.9, pch = 1)+
  scale_color_manual(name = "Station", values = c("red", "blue"))+
  facet_wrap(~year) # lovely!
```

## Zooming in on the south
```{r}
new_bbox <- st_bbox(feeding_cropped)
new_bbox[4] <- 31.501
new_bbox[2] <- 30.401

feeding_cropped_south <- st_crop(feeding_cropped, new_bbox)

#rosm::osm.types()
ggplot(feeding_cropped_south)+
  annotation_map_tile(zoom = 9, type = "hotstyle") +
  geom_sf(aes(col = station), alpha = 0.9, pch = 1)+
  scale_color_manual(name = "Station", values = c("red", "blue"))+
  facet_wrap(~year) # lovely!
```
## Can we use spatsoc to group the feeding bouts together into non-station carcass locations?
```{r}
ns_bouts <- feeding_cropped_south %>%
  filter(station == FALSE) %>%
  arrange(timestamp) %>%
  st_transform(32636)
ns_bouts$x <- st_coordinates(ns_bouts)[1,]
ns_bouts$y <- st_coordinates(ns_bouts)[2,]
dim(ns_bouts)

# Prepare data for analysis with spatsoc
ns_bouts$timestamp <- as.POSIXct(ns_bouts$timestamp)
ns_bouts <- data.table::data.table(ns_bouts)
class(ns_bouts$timestamp)

group_times(ns_bouts, datetime = 'timestamp', threshold = '24 hours') # group into 24 hour groups
# group_pts(ns_bouts, threshold = 0.1, id ='fake_id', coords = c('x', 'y'), timegroup = 'timegroup')
ns_bouts <- as.data.frame(ns_bouts)
ns_bouts <- st_as_sf(ns_bouts, crs = 32636, remove = F, sf_column_name = "geometry")
 #XXX WHAT IS GOING WRONG HERE?? THESE DISTANCES MAKE NO SENSE.

distMatrix <- as.matrix(stats::dist(cbind(ns_bouts$x, 
            ns_bouts$y), method = "euclidean"))
thresh <- 100 # units will be meters bc coordinates are in UTM
test <- distMatrix[1:100, 1:100]
test[test >= thresh] <- NA
g <- igraph::graph_from_adjacency_matrix(test, weighted = T, mode = "undirected")
plot(g)
graphAdj <- igraph::graph_from_adjacency_matrix(distMatrix)

DT[, `:=`(withinGroup, {
        distMatrix <- as.matrix(stats::dist(cbind(get(..coords[1]), 
            get(..coords[2])), method = "euclidean"))
        graphAdj <- igraph::graph_from_adjacency_matrix(distMatrix <= 
            threshold)
        igraph::clusters(graphAdj)$membership
    }), by = c(splitBy, timegroup), .SDcols = c(coords, id)]
    DT[, `:=`(group, .GRP), by = c(splitBy, timegroup, 
        "withinGroup")]

length(unique(ns_bouts$timegroup)) 
length(unique(ns_bouts$dateOnly)) # okay, we only have 67 days of data, so it makes sense that this is how many groups we ended up with.
length(unique(ns_bouts$group)) # more groups than days--this is as expected!

ns_bouts <- as.data.frame(ns_bouts) %>%
  st_as_sf(crs = 32636, remove = F) %>%
  rename("ns_carcass" = group)

mapview(ns_bouts %>% filter(ns_carcass == 3))
```

```{r}
ns_carcasses <- ns_bouts %>%
  group_by(year, ns_carcass) %>%
  summarize(geometry = st_union(geometry),
            dateOnly = dateOnly[1]) %>%
  st_centroid() %>%
  ungroup()
dim(ns_bouts)
dim(ns_carcasses)
head(ns_carcasses)
```

```{r}
ggplot()+
  annotation_map_tile(zoom = 9, type = "hotstyle") +
  geom_sf(data = ns_bouts, alpha = 0.5, pch = 1, col = "black")+
  geom_sf(data = ns_carcasses, alpha = 0.9, pch = 8, col = "red", size = 2)+
  facet_wrap(~year) # this is clearly not right! We expect the black and red points to be on top of each other.
```

## Visualizing station vs. non-station carcasses

## XXX insert carcass frequencies here (from report.qmd)